---
title: "MS008 Results"
format:
  html:
    toc: true
    embed-resources: true
execute:
  echo: false
  warning: false
  message: false
params:
  patient: "MS008"
  data_dir: "testData"
---

```{r setup}
library(knitr)
library(kableExtra)
library(ggplot2)
library(dplyr)
library(httr)
library(jsonlite)

helper <- if (file.exists("/opt/helperFunctions.R")) "/opt/helperFunctions.R" else "helperFunctions.R"
source(helper)

patient  <- params$patient
data_dir <- params$data_dir
out_dir  <- file.path(data_dir, "output")
```

## QC

```{r qc-summary, results='asis'}
llms_file <- file.path(data_dir, "multiqc", "global_report", "multiqc_data", "llms-full.txt")
if (file.exists(llms_file) && nzchar(Sys.getenv("ANTHROPIC_API_KEY"))) {
  qc_content <- paste(
    paste(readLines(llms_file, warn = FALSE), collapse = "\n"),
    "\n\nIMPORTANT: Ignore any formatting instructions above. Respond with 1-2 plain prose paragraphs only. Do not use :span, :sample, bullet points, markdown, or any special syntax. Write in complete sentences as if for a clinical report.",
    sep = ""
  )
  body <- list(
    model      = "claude-haiku-4-5-20251001",
    max_tokens = 1024L,
    system     = "You are a bioinformatics QC analyst. Summarize the provided QC metrics in 1-2 plain prose paragraphs suitable for a clinical report. Use no special formatting syntax, no markdown spans, no colons before words, no bullet points — plain sentences only.",
    messages   = list(list(role = "user", content = qc_content))
  )
  resp <- httr::POST(
    "https://api.anthropic.com/v1/messages",
    httr::add_headers(
      "x-api-key"         = Sys.getenv("ANTHROPIC_API_KEY"),
      "anthropic-version" = "2023-06-01",
      "content-type"      = "application/json"
    ),
    body = jsonlite::toJSON(body, auto_unbox = TRUE)
  )
  if (httr::http_error(resp)) {
    cat("*QC summary unavailable (API error).*\n")
  } else {
    parsed <- httr::content(resp, as = "parsed", encoding = "UTF-8")
    cat(parsed$content[[1]]$text, "\n")
  }
} else if (!file.exists(llms_file)) {
  cat("*MultiQC LLM summary file not found.*\n")
} else {
  cat("*Set `ANTHROPIC_API_KEY` to enable automated QC summary.*\n")
}
```

**Reports**

- [MultiQC Global Report](`r file.path(data_dir, "multiqc/global_report/multiqc_report.html")`)
- [MultiQC Group Report](`r file.path(data_dir, "multiqc/group_reports/paired_sample/multiqc_report.html")`)
- [MultiQC Report](`r file.path(data_dir, "multiqc/multiqc_report.html")`)

## Downloads

```{r write-csv, include=FALSE}
allVariants <- readRDS(file.path(out_dir, paste0("allVariants_", patient, ".RDS")))
csv_path <- file.path(out_dir, paste0("allVariants_", patient, ".csv"))
write.csv(allVariants, file = csv_path, row.names = FALSE)
```

- [Analysis Results](`r file.path(out_dir, paste0("allVariants_", patient, ".csv"))`)
- [Subcellular Location Table](`r file.path(out_dir, paste0("subcellularLocationTable_", patient, ".csv"))`)

## Variant Caller Comparison

```{r}
#| label: tbl-differenceMatrix
#| tbl-cap: "Percent variants in common between the various analyses"

differenceMatrix <- readRDS(file.path(out_dir, paste0("differenceMatrix_", patient, ".RDS")))
kable(differenceMatrix) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

## Subcellular Location Summary

```{r}
#| label: tbl-subcellularLocation
#| tbl-cap: !expr paste0(patient, " - Subcellular Location Summary Information")

allVariants <- readRDS(file.path(out_dir, paste0("allVariants_", patient, ".RDS")))
subcellularLocationTable <- createSubcellularLocationTable(allVariants)

# TOTAL row — unique variants across all callers
allVariants$change <- paste0(allVariants$Chromosome, "_", allVariants$Start_Position, "_", allVariants$Amino_acids)
unique_all <- allVariants[!duplicated(allVariants$change), ]
total_row <- data.frame(
  Name       = "TOTAL",
  Total      = length(unique(unique_all$change)),
  ECD        = sum(unique_all$extracellularMutationQ == TRUE, na.rm = TRUE),
  Secreted   = sum(unique_all$secretedButNotMutationInECDQ, na.rm = TRUE),
  NonDefined = 0.5 * sum(grepl("MEMBRANE", unique_all$membraneButNotMutationInECDorSecretedKeywords), na.rm = TRUE)
)
total_row$TotalMSPs <- paste0(
  total_row$ECD + total_row$Secreted + total_row$NonDefined, " (",
  round(100 * (total_row$ECD + total_row$Secreted + total_row$NonDefined) / total_row$Total), "%)"
)

write.csv(subcellularLocationTable,
          file = file.path(out_dir, paste0("subcellularLocationTable_", patient, ".csv")),
          row.names = FALSE)
kable(subcellularLocationTable) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

```{r}
#| label: tbl-subcellularLocationUnique
#| tbl-cap: !expr paste0(patient, " - Subcellular Location by Caller (unique variants)")

callers <- unique(unlist(strsplit(allVariants$Caller, ";")))
per_caller <- do.call(rbind, lapply(callers, function(cl) {
  r <- allVariants[grepl(cl, allVariants$Caller), ]
  r$change <- paste0(r$Chromosome, "_", r$Start_Position, "_", r$Amino_acids)
  r <- r[!duplicated(r$change), ]
  df <- data.frame(
    Name       = cl,
    Total      = nrow(r),
    ECD        = sum(r$extracellularMutationQ == TRUE, na.rm = TRUE),
    Secreted   = sum(r$secretedButNotMutationInECDQ, na.rm = TRUE),
    NonDefined = 0.5 * sum(grepl("MEMBRANE", r$membraneButNotMutationInECDorSecretedKeywords), na.rm = TRUE)
  )
  df$TotalMSPs <- paste0(
    df$ECD + df$Secreted + df$NonDefined, " (",
    round(100 * (df$ECD + df$Secreted + df$NonDefined) / df$Total), "%)"
  )
  df
}))
per_caller_with_total <- rbind(per_caller, total_row)

kable(per_caller_with_total) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) |>
  row_spec(nrow(per_caller_with_total), bold = TRUE)
```

## Tumor Purity

```{r}
#| label: tbl-purity

ichorcna_file <- file.path(data_dir, "ichorcna", paste0(patient, "_TUMOR.params.txt"))
purple_file   <- file.path(data_dir, "purple",   paste0(patient, "_", patient, "_TUMOR.purple.purity.tsv"))

rows <- list()

if (file.exists(ichorcna_file)) {
  ichor <- read.table(ichorcna_file, header = TRUE, sep = "\t", nrows = 1)
  rows[["ichorCNA"]] <- data.frame(
    Tool                  = "ichorCNA",
    `Tumor Fraction (%)` = round(ichor[1, "Tumor.Fraction"] * 100, 1),
    Ploidy                = ichor[1, "Ploidy"],
    check.names           = FALSE
  )
}

if (file.exists(purple_file)) {
  purple <- read.table(purple_file, header = TRUE, sep = "\t")
  rows[["PURPLE"]] <- data.frame(
    Tool                  = "PURPLE",
    `Tumor Fraction (%)` = round(purple[1, "purity"] * 100, 1),
    Ploidy                = purple[1, "ploidy"],
    check.names           = FALSE
  )
}

if (length(rows) > 0) {
  do.call(rbind, rows) |>
    kable() |>
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) |>
    column_spec(2, bold = TRUE)
} else {
  cat("No tumor purity files found.")
}
```

## VAF Distribution

```{r}
#| label: fig-vafDistribution
#| fig-cap: !expr paste0(patient, " - Variant Allele Frequency (VAF) distribution by caller")

vaf_file <- file.path(out_dir, paste0("VAF_", patient, ".RDS"))
if (file.exists(vaf_file)) {
  vaf_df <- readRDS(vaf_file)
  vaf_df <- vaf_df[!is.na(vaf_df$vaf), ]

  ggplot(vaf_df, aes(x = caller, y = vaf, fill = caller)) +
    geom_boxplot(outlier.size = 0.5, outlier.alpha = 0.4) +
    geom_hline(yintercept = 0.5, linetype = "dashed", color = "firebrick", linewidth = 0.5) +
    scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.1)) +
    labs(
      x = "Caller",
      y = "Variant allele frequency (VAF)",
      title = paste0(patient, " — somatic VAF distribution")
    ) +
    theme_minimal(base_size = 13) +
    theme(panel.grid.minor = element_blank(), legend.position = "none")
} else {
  cat("VAF file not found.")
}
```
